---
title: "Long Lapsed Donors"
author: "Ammar Ali"
date: "3/16/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

These donors gave in 2013 / 2014 but have not given since

```{r loading the data, include=FALSE}
data <- read_csv("Founders Club Prospects - Long Lapsed Prospects.csv")
colnames(data) <- c("Name", "Number.of.Donations", "Total.Amount", "Average.Amount")

```

## Step 1: Sorting by Number of donations

```{r Step 1.1, echo=FALSE}
hist1 <- ggplot(data, aes(x = Number.of.Donations)) +
  geom_histogram() +
  labs(title = "Distribution of Number of Donations") +
  xlab("Number of Donations") +
  ylab("Frequency")
hist1

```

Majority of the donors from this time period only gave one or two times. This next graph
look at the distribution of donations excluding one and two time donors

```{r Step 1.2, echo=FALSE}
data_1.2 <- data %>%
  filter(Number.of.Donations > 2)

hist2 <- ggplot(data_1.2, aes(x = Number.of.Donations)) +
  geom_histogram() +
  labs(title = "Distribution of Number of Donations", caption = "Filtered for donors with more than 2 donations.") +
  xlab("Number of Donations") +
  ylab("Frequency") +
  scale_x_continuous(breaks = round(seq(min(data_1.2$Number.of.Donations), max(data_1.2$Number.of.Donations), by = 2),1))
hist2
```

I will break down this dataset based on the number of donations: more than 2 donations, 2 donations, and 1 donation

## Step 2: Donors with more than 2 donations.

```{r step 2, echo=FALSE}
DT::datatable(data_1.2)
```
A lot of these donors gave quite frequently in smaller amounts or a few times but with larger amounts. These seem like good founders club prospects just based on their intial enthusiam towards emgage.

## Step 3: Donors with 2 donations

```{r step 3, echo=FALSE}
data_3 <- data %>%
  filter(Number.of.Donations == 2) %>%
  arrange(desc(Total.Amount))
DT::datatable(data_3)

```
First 30 rows in this table look like good potential donors. Rows 1 -3  have Founders club potential.

## Step 4: One Time Donors

```{r step 4, echo=FALSE}
data_4 <- data %>%
  filter(Number.of.Donations == 1) %>%    
  arrange(desc(Total.Amount))
data_4 <- data_4[, -4]
DT::datatable(data_4)

hist_3 <- ggplot(data_4, aes(Total.Amount)) +
  geom_histogram() +
  labs(title = "Distribution of Total Donation Amounts") +
  xlab("Total Donation Amount") +
  ylab("Frequency")
hist_3
```


The first 51 donors from this category gave a one time donation greater than or equal to 1000 